---
title: "Interactive Visualizations"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---


Below are some of the visualizations created with this project.

```{r message=FALSE, echo=FALSE, include=FALSE, warning=FALSE}

library(tidyverse)
library(plotly)
library(widgetframe)
library(tidytext)

library(data.table)
library(kableExtra)
library(dplyr)
library(splines)
library(mgcv)
library(gridExtra)
library(grid)
library(cowplot)
library(broom)
```


```{r, echo=FALSE}
fn1 <- "https://raw.githubusercontent.com/inorrr/JSC370_project/main/census.csv"
fn2 <- "https://raw.githubusercontent.com/inorrr/JSC370_project/main/crime.csv"

if (!file.exists("data/census.csv"))
  download.file(fn1, destfile = "data/census.csv")
census_df <- data.table::fread("data/census.csv")

if (!file.exists("data/crime.csv"))
  download.file(fn2, destfile = "data/crime.csv")
crime_df <- data.table::fread("data/crime.csv")

crime_df <- pivot_wider(crime_df, id_cols = c(REF_DATE, GEO, Violations), 
                        names_from = Statistics, values_from = VALUE)
crime_df <- na.omit(crime_df)

census_df <- pivot_wider(census_df, id_cols = c(REF_DATE, GEO, `Income source`), 
                         names_from = Statistics, values_from = VALUE)
census_df <- na.omit(census_df)
```

```{r, echo=FALSE}
unique_x_values <- unique(crime_df$REF_DATE)

filtered_data = crime_df %>% filter(Violations=="Total, all violations")
plot1 <- ggplot(filtered_data, aes(x = REF_DATE, y = `Rate per 100,000 population`, color = GEO)) +
  geom_line() +
  labs(x = "Year", y = "Rate per 100,000 population", title = "Rates of Total Crime by Province") +
  scale_x_continuous(breaks = unique_x_values) + 
  scale_color_discrete(name = "Provinces") + 
  theme(legend.position = "bottom") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(plot.title = element_text(face = "bold")) + 
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")) + 
  theme(axis.text = element_text(size = 7)) + 
  theme(legend.title = element_text(face = "bold"))


filtered_data = crime_df %>% filter(Violations=="Total breaking and entering")
plot2 <- ggplot(filtered_data, aes(x = REF_DATE, y = `Rate per 100,000 population`, color = GEO)) +
  geom_line() +
  labs(x = "Year", y = "Rate per 100,000 population", title = "Rates of Break and Enter by Province") +
  scale_x_continuous(breaks = unique_x_values) + 
  scale_color_discrete(name = "Provinces") + 
  theme(legend.position = "bottom") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(plot.title = element_text(face = "bold")) + 
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")) + 
  theme(axis.text = element_text(size = 7)) + 
  theme(legend.title = element_text(face = "bold"))

filtered_data = crime_df %>% filter(Violations=="Total robbery")
plot3 <- ggplot(filtered_data, aes(x = REF_DATE, y = `Rate per 100,000 population`, color = GEO)) +
  geom_line() +
  labs(x = "Year", y = "Rate per 100,000 population", title = "Rates of Robbery by Province") +
  scale_x_continuous(breaks = unique_x_values) + 
  scale_color_discrete(name = "Provinces") + 
  theme(legend.position = "bottom") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(plot.title = element_text(face = "bold")) + 
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")) + 
  theme(axis.text = element_text(size = 7)) + 
  theme(legend.title = element_text(face = "bold"))

filtered_data = crime_df %>% filter(Violations=="Total prostitution")
plot4 <- ggplot(filtered_data, aes(x = REF_DATE, y = `Rate per 100,000 population`, color = GEO)) +
  geom_line() +
  labs(x = "Year", y = "Rate per 100,000 population", title = "Rates of Prostitution by Province") +
  scale_x_continuous(breaks = unique_x_values) + 
  scale_color_discrete(name = "Provinces") + 
  theme(legend.position = "bottom") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(plot.title = element_text(face = "bold")) + 
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")) + 
  theme(axis.text = element_text(size = 7)) + 
  theme(legend.title = element_text(face = "bold"))
```


# Crime Rates By Province {.tabset}

The line plots shows the crime rates by provinces for each year.

## Total Crime

```{r plot_1, echo=FALSE}
ggplotly(plot1, width = 800)
```

## Break and Enter

```{r plot_2, echo=FALSE}
ggplotly(plot2, width = 800)
```

## Robbery

```{r plot_3, echo=FALSE}
ggplotly(plot3, width = 800)
```

## Prostitution

```{r plot_4, echo=FALSE}
ggplotly(plot4, width = 800)
```

# Average Income by Province {.tabset}

The line plots shows the average income of provinces each year, this can help with identifying the trend of income. 

```{r, echo = FALSE, fig.align='center', fig.width=10, fig.height=5}
filtered_data = census_df %>% filter(`Income source`=="Total income")
unique_x_values <- unique(census_df$REF_DATE)
plot1 <- ggplot(filtered_data, aes(x = REF_DATE, y = `Average income (excluding zeros)`, color = GEO)) +
  geom_line() +
  labs(x = "Year", y = "Total income", title = "Average Total Income (excluding zeros) by Province") +
  scale_x_continuous(breaks = unique_x_values) +
  scale_color_discrete(name = "Provinces") +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")) + 
  theme(axis.text = element_text(size = 7)) +
  theme(legend.title = element_text(face = "bold"))


filtered_data = census_df %>% filter(`Income source`=="Employment income")
plot2 <- ggplot(filtered_data, aes(x = REF_DATE, y = `Average income (excluding zeros)`, color = GEO)) +
  geom_line() +
  labs(x = "Year", y = "Average Income (excluding zeros)", title = "Average Employment Income by Province") +
  scale_x_continuous(breaks = unique_x_values) +
  scale_color_discrete(name = "Provinces") +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")) + 
  theme(axis.text = element_text(size = 7)) +
  theme(legend.title = element_text(face = "bold"))

filtered_data = census_df %>% filter(`Income source`=="Investment income")
plot3 <- ggplot(filtered_data, aes(x = REF_DATE, y = `Average income (excluding zeros)`, color = GEO)) +
  geom_line() +
  labs(x = "Year", y = "Average Income (excluding zeros)", title = "Average Investment Income by Province") +
  scale_x_continuous(breaks = unique_x_values) +
  scale_color_discrete(name = "Provinces") +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")) + 
  theme(axis.text = element_text(size = 7)) +
  theme(legend.title = element_text(face = "bold"))

filtered_data = census_df %>% filter(`Income source`=="Market income")
plot4 <- ggplot(filtered_data, aes(x = REF_DATE, y = `Average income (excluding zeros)`, color = GEO)) +
  geom_line() +
  labs(x = "Year", y = "Average Income (excluding zeros)", title = "Average Market Income by Province") +
  scale_x_continuous(breaks = unique_x_values) +
  scale_color_discrete(name = "Provinces") +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")) + 
  theme(axis.text = element_text(size = 7)) +
  theme(legend.title = element_text(face = "bold"))
```


## Total Income
```{r echo=FALSE}
ggplotly(plot1, width = 800)
```

## Employment Income
```{r echo=FALSE}
ggplotly(plot2, width = 800)
```

## Investment Income
```{r echo=FALSE}
ggplotly(plot3, width = 800)
```

## Market Income
```{r echo=FALSE}
ggplotly(plot4, width = 800)
```

# Trend in Average Total Income

The plots shows the average total income of all provinces in each year. Thus we can identify overall trends of income by year.

```{r, echo=FALSE, fig.align='center', fig.width=10, fig.height=5}
filtered_data = census_df %>% filter(`Income source`=="Total income")
plot<-ggplot(filtered_data, mapping = aes(x = as.factor(REF_DATE), y = `Average income (excluding zeros)`)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.8, color = "pink") + 
  labs(x = "Year", y = "Total income", title = "Average Total Income (excluding zeros)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")) + 
  theme(axis.text = element_text(size = 7))

ggplotly(plot, width = 800)

```


# Crime Rate vs Income{.tabset}

The scatter plot shows the relationshio between crime rate and income. While the first tab shows total crime(all types) and total income, the rest demonstrates the relationship between a specific crime type and income type. 


```{r, echo = FALSE, warning = FALSE, fig.align='center', fig.width=10, fig.height=5}
filtered_census <- census_df %>% filter(`Income source` == "Total income") 
filtered_crime <- crime_df %>%filter(Violations=="Total, all violations")

joint_df <- inner_join(filtered_crime, filtered_census, by = c("REF_DATE", "GEO"))


plot1 <- ggplot(data = joint_df, aes(x = `Average income (excluding zeros)`, y = `Rate per 100,000 population`, color = GEO, size = 3)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE, size = 1.0) + 
  guides(size = FALSE) + # remove "size" from legend
  labs(x = "Average Total Income (excluding zeros)", 
       y = "Total Crime Rate per 100,000 population`", 
       title = "Average Total Income and Total Crime Rate") +
  scale_color_discrete(name = "Provinces") +
  theme(axis.text = element_text(size = 7)) +
  theme(plot.title = element_text(face = "bold")) +
  theme(legend.position = "right") +
  theme(legend.title = element_text(face = "bold")) +
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"))

filtered_census <- census_df %>% filter(`Income source` == "Employment income") 
filtered_crime <- crime_df %>%filter(Violations=="Total robbery")

joint_df <- inner_join(filtered_crime, filtered_census, by = c("REF_DATE", "GEO"))

plot2 <- ggplot(data = joint_df, aes(x = `Average income (excluding zeros)`, y = `Rate per 100,000 population`, color = GEO, size = 3)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE, size = 1.0) + 
  guides(size = FALSE) + # remove "size" from legend
  labs(x = "Average Employment Income (excluding zeros)", 
       y = "Robbery Crime Rate per 100,000 population`", 
       title = "Avg Employment Income vs Robbery Crime Rate") +
  scale_color_discrete(name = "Provinces") +
  theme(axis.text = element_text(size = 7)) +
  theme(plot.title = element_text(face = "bold")) +
  theme(legend.position = "right") +
  theme(legend.title = element_text(face = "bold")) +
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"))

filtered_census <- census_df %>% filter(`Income source` == "Market income") 
filtered_crime <- crime_df %>%filter(Violations=="Total property crime violations")

joint_df <- inner_join(filtered_crime, filtered_census, by = c("REF_DATE", "GEO"))

plot3 <- ggplot(data = joint_df, aes(x = `Average income (excluding zeros)`, y = `Rate per 100,000 population`, color = GEO, size = 3)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE, size = 1.0) + 
  guides(size = FALSE) + # remove "size" from legend
  labs(x = "Average Market Income (excluding zeros)", 
       y = "Property Crime Rate per 100,000 population`", 
       title = "Avg Market Income vs Property Crime Rate") +
  scale_color_discrete(name = "Provinces") +
  theme(axis.text = element_text(size = 7)) +
  theme(plot.title = element_text(face = "bold")) +
  theme(legend.position = "right") +
  theme(legend.title = element_text(face = "bold")) +
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"))

filtered_census <- census_df %>% filter(`Income source` == "Government transfers") 
filtered_crime <- crime_df %>%filter(Violations=="Total prostitution")

joint_df <- inner_join(filtered_crime, filtered_census, by = c("REF_DATE", "GEO"))

plot4 <- ggplot(data = joint_df, aes(x = `Average income (excluding zeros)`, y = `Rate per 100,000 population`, color = GEO, size = 3)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE, size = 1.0) + 
  guides(size = FALSE) + # remove "size" from legend
  labs(x = "Average Employment Income (excluding zeros)", 
       y = "Prostitution Crime Rate per 100,000 population`", 
       title = "Avg Employment Income vs Prostitution Crime Rate") +
  scale_color_discrete(name = "Provinces") +
  theme(axis.text = element_text(size = 7)) +
  theme(plot.title = element_text(face = "bold")) +
  theme(legend.position = "right") +
  theme(legend.title = element_text(face = "bold")) +
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"))
```

## Total Income and Total Crime
```{r echo=FALSE, }
ggplotly(plot1, width = 800)
```

## Employment Income and Robbery
```{r echo=FALSE}
ggplotly(plot2, width = 800)
```

## Market Income and Property Crime
```{r echo=FALSE}
ggplotly(plot3, width = 800)
```

## Employment Income and Prostitution
```{r echo=FALSE}
ggplotly(plot4, width = 800)
```

# Income Levels

I created 4 levels for average total income using the quarantines, from negative infinity to the first quantile is **"Low"**, from first quantile to mean is **"Med_Low"**, from mean to 3rd quantile is **"Med_High"**, from 3rd quantile above is **"High"**. The box plot shows the average total crime rate by income level.

```{r, echo = FALSE}
filter_census <- census_df %>% filter(`Income source` == "Total income")
# summary(filter_census$`Average income (excluding zeros)`)
breaks <- c(-Inf, 40460, 45077, 48859, Inf)
filter_census$income_level <- cut(filter_census$`Average income (excluding zeros)`, 
                                  breaks = breaks, labels = c("Low", "Med-Low", "Med-High", "High"))


filter_crime <- crime_df %>%filter(Violations=="Total, all violations")
joint_df <- inner_join(filter_census, filter_crime, by = c("REF_DATE", "GEO"))

plot<-ggplot(joint_df, aes(x = as.factor(income_level), y = `Rate per 100,000 population`)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.8, color = "tomato") + 
  labs(x = "Income Level", y = "Total Crime Rate", title = "Total Crime Rate by Income Level") +
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")) +
  theme(axis.text = element_text(size = 10))

ggplotly(plot, width = 800)
```


